---
title: "Equally-Weighted Risk Contributions by optimization "
author: "Richard Warnung"
date: "July 8, 2022"
---
# Introduction
The approach is based on [On the Properties of Equally-Weighted Risk Contributions Portfolios](https://www.researchgate.net/publication/45397778_On_the_Properties_of_Equally-Weighted_Risk_Contributions_Portfolios)
by Teiletche and Roncalli. \
Required packages: alabama 

# Generating data
4 assets with decreasing volatility and equal correlation:

```{r }

vols = matrix( c(0.3, 0.25, 0.2, 0.15), nrow = 4)

rho = 0.7
cor.mat = matrix(c(rho), nrow=4, ncol=4)
diag(cor.mat) = 1

sigma.mat =  cor.mat * matrix(outer(vols, vols), nrow=4, byrow=TRUE) 

sigma.mat
## checking the covariance matrix
cov2cor(sigma.mat)
```

# Equally weighted portfolio

```{r }
x_equal = matrix(c(1/4, 1/4, 1/4, 1/4), nrow=4)
```

volatility and contributions of equally weighted portfolio:

```{r }
vol_equal = sqrt( t(x_equal)%*%sigma.mat%*%x_equal )[1,1]
vol_equal
contribs_equal_w = x_equal *(sigma.mat%*%x_equal/vol_equal)
contribs_equal_w
```

# Equally-Weighted Risk Contributions (ERC) portfolio
Optimization with `constrOptim.nl` from package `alabama`.
We first solve the problem
$$\sqrt{y^T \Sigma y} \rightarrow Min$$ 
subject to
\begin{align*}
   y_i > 0 \\
   \sum \ln y_i \ge c
\end{align*} 
Then we rescale $x_i = y_i/\sum_{i=1}^n y_i$ to get the ERC portfolio.


```{r }
library(alabama)

objective_fn = function(par){
  return( (t(par)%*%sigma.mat%*%par)[1,1]) ## no square-root, same arg min
}

c_max = -nrow(sigma.mat)*log(nrow(sigma.mat))

constr_fn = function(par){
  h = rep(NA, length(par)+1)
  h[1:length(par)] = par # par[i] > 0
  h[length(par)+1] =  sum(log(par))-c  # sum(log(par)) > c
  return( h )
}

oldw = getOption("warn") ## silencing the warnings during optimization, keeping the settings for later.
options(warn = -1)
```

## Choosing c = c_max-0.4

```{r }
c = c_max-0.4

ans = constrOptim.nl(par = c(1/5, 1/5, 1/5, 1-3/5), fn = objective_fn, hin = constr_fn, control.outer = list(trace = FALSE))  
y = ans$par
y
sum(y)
x = y/sum(y)
x
vol_ERC = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_ERC

contribs = x *(sigma.mat%*%x/vol_ERC)
contribs
```

## Choosing c = c_max-0.5

```{r }
c = c_max-0.5

ans = constrOptim.nl(par = c(1/5, 1/5, 1/5, 1-3/5), fn = objective_fn, hin = constr_fn,  control.outer = list(trace = FALSE))  
y2 = ans$par
y2
sum(y2)
x = y2/sum(y2)
x
vol_ERC = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_ERC
contribs = x *(sigma.mat%*%x/vol_ERC)
contribs
```

## Choosing c = c_max-1

```{r }
c = c_max-1

ans = constrOptim.nl(par = c(1/5, 1/5, 1/5, 1-3/5), fn = objective_fn, hin = constr_fn,  control.outer = list(trace = FALSE))  
y3 = ans$par
y3
sum(y3)
x = y3/sum(y3)
x
vol_ERC = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_ERC
contribs = x *(sigma.mat%*%x/vol_ERC)
contribs
```

## Conclusion

As long as $c$ is below the upper bound $-n \log(n)$, solutions $y$ are different but lead to the ERC after scaling.

# Direct derivation of the ERC portfolio without scaling
This time we solve the following optimization problem: 
$$\sqrt{x^T \Sigma x} \rightarrow Min$$ 
subject to
\begin{align*}
   x_i &> 0 \\
   \sum_{i=1}^n x_i &= 1 \\
   \sum_{i=1}^n \ln x_i &\ge c^*.
\end{align*}

We define a new constraint function and fix the constant to $c^{\ast} = c - n \ln( \sum_{i=1}^n y_i^{\ast})$. Where $c$ was used to derive the optimal solution $y^{\ast}$.
According to Teiletche and Roncalli this will result in the ECR directly without the need to rescale.

```{r }
c_star = c - length(y3)*log(sum(y3))
constr_fn_star = function(par){
  h = rep(NA, length(par)+1)
  h[1:length(par)] = par # par[i] > 0
  h[length(par)+1] =  sum(log(par))-c_star  # sum(log(par)) > c*
  return( h )
}

constr_eq_fn = function(par){
  h = rep(NA,1)
  h[1] = sum(par)-1 
}

ans = constrOptim.nl(par = c(0.18, 0.22, 0.29, 0.31), fn = objective_fn, hin = constr_fn_star, heq = constr_eq_fn, control.outer = list(trace = FALSE))  

x = ans$par
x
sum(x)

vol_ERC = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_ERC
contribs = x *(sigma.mat%*%x/vol_ERC)
contribs
```

We see that the solution $x$ gives the ERC (apart from numerical inaccuracies).

## Changing the constant $c^*$

Next we investigate what happens if we change the constant c_star. We go back to the initial constant $c^* = c$.

```{r }

c_star = c 
ans = constrOptim.nl(par = c(0.18, 0.22, 0.29, 0.31), fn = objective_fn, hin = constr_fn_star, heq = constr_eq_fn, control.outer = list(trace = FALSE))  

x = ans$par
x
sum(x)

vol_pf = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_pf
contribs = x *(sigma.mat%*%x/vol_pf)
contribs
```

We see that the portfolio constraint is still fulfilled, but the ERC-property is not fulfilled. The volatility is lower than in the case of the ERC.

## Removing the portfolio constraint

Next we investigate the result if we remove the portfolio constraint $\sum x_i = 1$:

```{r }
c_star = c - length(y3)*log(sum(y3))
ans = constrOptim.nl(par = c(0.18, 0.22, 0.29, 0.31), fn = objective_fn, hin = constr_fn_star, control.outer = list(trace = FALSE))  

x = ans$par
x
sum(x)

vol_pf = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_pf
contribs = x *(sigma.mat%*%x/vol_pf)
contribs
```

The ERC and the portfolio property are fulfilled.


## Changing the objective function

Due to the dominance of the condition $\sum_{i=1}^n \ln x_i \ge c^{\ast}$ we investigate whether we can replace the objective function by another convex function. We choose
the L2-norm $\sum_{i=1}^n x_i^2$.


```{r }
```
```{r }

objective_fn = function(par){
  return(  sum(par^2) ) ## norm as objective function
}

ans = constrOptim.nl(par = c(0.18, 0.22, 0.29, 0.31), fn = objective_fn, hin = constr_fn_star, control.outer = list(trace = FALSE))  

x = ans$par
x
sum(x)

vol_pf = sqrt( t(x)%*%sigma.mat%*%x)[1,1]
vol_pf
contribs = x *(sigma.mat%*%x/vol_pf)
contribs
```

Changing the objective function destroys the ERC property.

## Derivation of the mathematics

Let us see why $c^{\ast} = c - n \ln( \sum_{i=1}^n y_i^{\ast})$ gives the correct solution.
Looking at the constraint $\sum_{i=1}^n \ln x_i \ge c^{\ast}$ it follows due to the Kuhn–Tucker conditions that the inequality is actually reached:
$$
 \sum_{i=1}^n \ln x_i = c^*
$$
Then we insert for $c^*$ and due to the choice of $c$ we can also insert $c = \sum_{i=1}^n \ln(y_i^{\ast})$:
\begin{align*}
 \sum_{i=1}^n \ln x_i &= c - n \ln( \sum_{i=1}^n y_i^{\ast}) \\
 \sum_{i=1}^n \ln x_i &= \sum_{i=1}^n \ln(y_i^{\ast}) - n \ln( \sum_{i=1}^n y_i^{\ast})  \\
 \sum_{i=1}^n \ln x_i &= \sum_{i=1}^n \ln \left( \frac{y_i^{\ast}}{\sum_{i=1}^n y_i^{\ast}} \right).
\end{align*}


```{r include=FALSE}
options(warn = oldw)
```

